HaPy \\ Haskell for Python

### 2011-12-17 paper.tex
Added stuff about future work (converting ADTs to Python types) to paper

\section{Use}
Importing a Haskell module into a Python program is easy with HaPy: simply add the line
\begin{minted}{python}
from HaPy import HsModule
\end{minted}
to the head of any Python script. This will dynamically load the specified Haskell module, so that any Haskell function defined in HsModule can be accessed and executed by the Python script. \\
Installing HaPy is similarly intuitive, but with the caveat that all libraries called by HaPy.hs must be dynamically-linkable. This means that all of the packages that HaPy.hs imports must be reinstalled with Cabal, using the \verb!--enable-shared! flag. \footnote{The requirement that these packages be dynamically-linkable is enforced by Python's ctypes module, which we use to marshall primitive types across the Haskell-Python interface.} \\
However, Haskell packages that are used by the user's Haskell code need not be dynamically-linkable. Indeed, on the Haskell side, no special support is needed by the user; literally any standard Haskell code will run with HaPy. All the issues of marshalling data and passing data to and from the Foreign Function Interface are managed by the code in HaPy.hs \\
This has the added benefit of enabling the user to call any code defined in the Haskell standard libraries from Python. 
% - Meshing python and haskell semantics

\subsection{Loading Haskell Modules}
% - super easy: all haskell modules represented as submodules of HaPy
%  - standard haskell module represnetnation (e.x. System.Plugins) preserved
% - import HaPy.Data.List

@@

\subsection{Supporting more types}

% marshalling lists
% mashralling ADTs

>>>>

\section{Use}
Our goal with HaPy is to allow the user to call arbitrary Haskell code while remaining committed to ``Pythonic'' syntax. In particular, importing a Haskell module into a Python program is easy with HaPy: simply add the line
\begin{minted}{python}
from HaPy import HsModule
\end{minted}
or
\beign{minted}{python}
import HaPy.HsModule
\end{minted}
to the head of any Python script. This will dynamically load the specified Haskell module, so that any Haskell function defined in HsModule can be accessed and executed by the Python script, as though it were code defined in another Python module. For instance, if HsModule defines a function \verb!sum! which takes two real numbers and returns their sum, the result can be printed in Python with
\begin{minted}{python}
print HsModule.sum(first, second)
\end{minted}
Installing HaPy is similarly intuitive, but with the caveat that all libraries called by HaPy.hs must be dynamically-linkable. This means that all of the packages that HaPy.hs imports must be reinstalled with Cabal, using the \verb!--enable-shared! flag. \footnote{The requirement that these packages be dynamically-linkable is enforced by Python's ctypes module, which we use to marshall primitive types across the Haskell-Python interface.} \\
However, Haskell packages that are used by the user's Haskell code need not be dynamically-linkable. Indeed, on the Haskell side, no special support is needed by the user; literally any standard Haskell code will run with HaPy. All the issues of marshalling data and passing data to and from the Foreign Function Interface are managed by the code in HaPy.hs \\
This has the added benefit of enabling the user to call any code defined in the Haskell standard libraries from Python. 
% - Meshing python and haskell semantics

\subsection{Loading Haskell Modules}
As noted previously, the syntax for loading Haskell modules is identical to the normal Python import syntax. All Haskell modules are represented as submodules of the HaPy module, in order that the Python interpreter should not mistake them for Python modules. \\
Furthermore, the Haskell module hierarchy is preserved, so for instance one can say
\begin{minted}{python}
import HaPy.Data.List
\end{minted}
% - super easy: all haskell modules represented as submodules of HaPy
%  - standard haskell module represnetnation (e.x. System.Plugins) preserved
% - import HaPy.Data.List

@@

\subsection{Supporting more types}
In addition to marshalling primitive types across the Haskell-Python interface, we would eventually like to be able to marshall Python lists into Haskell lists and vice versa. One potential challenge here would be the marshalling of infinitely long, lazily-evaluated lists from Haskell to Python. \\
A bigger challenge would be dynamically converting from algebraic data types (ADTs) defined in Haskell to Python objects; for example, a simple product type composed of Haskell primitives could be represented as a named tuple\footnote{Effectively a fixed-size dictionary} in Python.
% marshalling lists
% mashralling ADTs
